\documentclass[conference]{IEEEtran}

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}

\title{Quantized Retrieval-Augmented Medical Vision-Language Diagnosis}

\author{
  \IEEEauthorblockN{Author Name}
  \IEEEauthorblockA{Institution\\Email}
}

\begin{document}
\maketitle

\begin{abstract}
This paper proposes a multimodal medical diagnosis pipeline that aligns medical images and clinical reports in a shared latent space, compresses the latent representation for efficient retrieval, and augments diagnostic reasoning with similar medical cases.
\end{abstract}

\begin{IEEEkeywords}
medical vision-language model, retrieval-augmented generation, quantization, FAISS, explainable diagnosis
\end{IEEEkeywords}

\section{Introduction}
Medical diagnosis from imaging data benefits from models that jointly reason over visual findings, historical reports, and related cases. We study a quantized retrieval-augmented vision-language pipeline for MRI, CT, and histopathology diagnosis.

\section{Related Work}
\subsection{Medical Vision-Language Learning}
Review biomedical image-text representation learning, contrastive alignment, and clinical report modeling.

\subsection{Retrieval-Augmented Medical Diagnosis}
Discuss similar-case retrieval, case-based reasoning, and retrieval-augmented generation for clinical decision support.

\subsection{Quantized Vector Retrieval}
Discuss product quantization, vector quantization, and compressed nearest-neighbor search.

\section{Method}
\subsection{Image and Text Encoders}
The image encoder maps MRI, CT, or histopathology inputs into visual embeddings. The text encoder maps clinical reports into biomedical text embeddings.

\subsection{Cross-Modal Alignment}
Image and report embeddings are projected into a shared latent space and optimized with a symmetric InfoNCE objective:
\begin{equation}
\mathcal{L} =
- \log
\frac{
\exp(\mathrm{sim}(z_i, z_t) / \tau)
}{
\sum_j \exp(\mathrm{sim}(z_i, z_j) / \tau)
}.
\end{equation}

\subsection{Quantized Retrieval}
The aligned medical embeddings are compressed before indexing. At inference time, a query image embedding retrieves similar cases from a vector database.

\subsection{Diagnosis and Report Generation}
Retrieved cases provide context for disease classification and report generation.

\section{Experiments}
\subsection{Datasets}
Candidate datasets include BraTS, FastMRI, IXI, MIMIC-CXR, CheXpert, CAMELYON, and PANDA.

\subsection{Metrics}
Evaluate classification accuracy, retrieval recall, report quality, calibration, latency, and compression ratio.

\section{Discussion}
Discuss explainability, uncertainty estimation, privacy, edge deployment, and clinical limitations.

\section{Conclusion}
We outline a research direction for efficient and explainable medical vision-language diagnosis through quantized retrieval-augmented reasoning.

\nocite{*}
\bibliographystyle{IEEEtran}
\bibliography{../paper-refs/references}

\end{document}
